---
title: "Homework_09"
author: "Benjamin Recchia"
date: "2024-05-26"
output: html_document
---

# #1: Structured Programming Version of Homework 7
```{r, message=FALSE,warning=FALSE}
# FUNCTIONS --------------------

#######
# FUNCTION: gen_data
# generate random data
# input: parameters n, lecturemean, lecturesd, labmean, labsd
# output: data frame of lecture and lab exam data with an ID column
#------------------
gen_data <- function(n=100,
                     lecturemean=80,
                     lecturesd=10,
                     labmean=40,
                     labsd=5) {
  lecture_exam <- rnorm(n=n,mean=lecturemean,sd=lecturesd)
  #generate simulated lecture exam data (hypothetically independent variable)
  lab_exam <- lecture_exam/2+rnorm(n=n,mean=lecturemean/2-labmean,sd=labsd)
  #generate simulated lab exam data (hypothetically dependent variable) from lecture exam data. Lab exams are half as many points as lecture exams, and I introduced some random normal variation based on the sd of the lab exam scores.
  ID <- 1:n #identifier column
  data_frame <- data.frame(ID,lecture_exam,lab_exam) #generate data frame
  return(data_frame)
  }

#######
# FUNCTION: calculate_linreg
# fit an ordinary least squares regression
# input: x and y vectors of same length
# output: entire summary of regression model
#------------------
calculate_linreg <- function(x_var=runif(10),
                            y_var=runif(10)) {
  data_frame <- data.frame(x_var,y_var)
  reg_model <- lm(y_var ~ x_var)
  return(unlist(summary(reg_model)))
}

#######
# FUNCTION: summarize_output
# pull elements from model summary list
# input: list from summary call of lm
# output: vector of regression residuals
#------------------
summarize_output <- function(z=NULL) {
  if(is.null(z)) {
    z <- summary(lm(runif(10)~runif(10)))
  }
    return(z$coefficients8)
}

#######
# FUNCTION: graph_results
# one line description
# input: x
# output: x
#------------------
graph_results <- function(x_var=runif(10),
                          y_var=runif(10)) {
  data_frame <- data.frame(x_var,y_var)
p1 <- ggplot2::ggplot(data_frame) + aes(x=x_var,y=y_var) + geom_point() + stat_smooth(method="lm")
  suppressMessages(print(p1))
}

# PRELIMINARY CALLS --------------------

library(tidyverse)
library(ggplot2)
set.seed(99)

# GLOBAL VARIABLES --------------------

n <- 40 #enrollment in the course
lecturemean <- 80 #mean of four lecture exam scores in course
lecturesd <- 10 #sd of four lecture exam scores in course
labmean <- 40 #mean of four lab exam scores in course
labsd <- 5 #sd of four lab exam scores in course

# PROGRAM BODY --------------------

temp_1 <- gen_data(n=n,
                   lecturemean = lecturemean,
                   lecturesd = lecturesd,
                   labmean = labmean,
                   labsd = labsd) #generate data frame from global variables

x <- temp_1$lecture_exam #extract x from data frame
y <- temp_1$lab_exam #extract y from data frame

temp_2 <- calculate_linreg(x_var=x,y_var=y) #perform linear regression

pval <- summarize_output(temp_2) #extract p value from linear regression stats
pval

plot1 <- graph_results(x_var=x, y_var=y) #plot data and linear regression
```

# #2: Same as #1 but returns R Squared instead of p Value
```{r, message=FALSE,warning=FALSE}
# FUNCTIONS --------------------

#######
# FUNCTION: gen_data
# generate random data
# input: parameters n, lecturemean, lecturesd, labmean, labsd
# output: data frame of lecture and lab exam data with an ID column
#------------------
gen_data <- function(n=100,
                     lecturemean=80,
                     lecturesd=10,
                     labmean=40,
                     labsd=5) {
  lecture_exam <- rnorm(n=n,mean=lecturemean,sd=lecturesd)
  #generate simulated lecture exam data (hypothetically independent variable)
  lab_exam <- lecture_exam/2+rnorm(n=n,mean=lecturemean/2-labmean,sd=labsd)
  #generate simulated lab exam data (hypothetically dependent variable) from lecture exam data. Lab exams are half as many points as lecture exams, and I introduced some random normal variation based on the sd of the lab exam scores.
  ID <- 1:n #identifier column
  data_frame <- data.frame(ID,lecture_exam,lab_exam) #generate data frame
  return(data_frame)
  }

#######
# FUNCTION: calculate_linreg
# fit an ordinary least squares regression
# input: x and y vectors of same length
# output: entire summary of regression model
#------------------
calculate_linreg <- function(x_var=runif(10),
                            y_var=runif(10)) {
  data_frame <- data.frame(x_var,y_var)
  reg_model <- lm(y_var ~ x_var)
  return(unlist(summary(reg_model)))
}

#######
# FUNCTION: summarize_output_r2
# pull elements from model summary list
# input: list from summary call of lm
# output: vector of regression residuals
#------------------
summarize_output_r2 <- function(z=NULL) {
  if(is.null(z)) {
    z <- summary(lm(runif(10)~runif(10)))
  }
    return(z$r.squared) #changed to R squared from p value
}

#######
# FUNCTION: graph_results
# one line description
# input: x
# output: x
#------------------
graph_results <- function(x_var=runif(10),
                          y_var=runif(10)) {
  data_frame <- data.frame(x_var,y_var)
p1 <- ggplot2::ggplot(data_frame) + aes(x=x_var,y=y_var) + geom_point() + stat_smooth(method="lm")
  suppressMessages(print(p1))
}

# PRELIMINARY CALLS --------------------

library(tidyverse)
library(ggplot2)
set.seed(99)

# GLOBAL VARIABLES --------------------

n <- 40 #enrollment in the course
lecturemean <- 80 #mean of four lecture exam scores in course
lecturesd <- 10 #sd of four lecture exam scores in course
labmean <- 40 #mean of four lab exam scores in course
labsd <- 5 #sd of four lab exam scores in course

# PROGRAM BODY --------------------

temp_1 <- gen_data(n=n,
                   lecturemean = lecturemean,
                   lecturesd = lecturesd,
                   labmean = labmean,
                   labsd = labsd) #generate data frame from global variables

x <- temp_1$lecture_exam #extract x from data frame
y <- temp_1$lab_exam #extract y from data frame

temp_2 <- calculate_linreg(x_var=x,y_var=y) #perform linear regression

r2 <- summarize_output_r2(temp_2) #extract R squared from linear regression stats
r2

plot1 <- graph_results(x_var=x, y_var=y) #plot data and linear regression
```

# #3:

smallest sample size that achieves >95% likelihood of >95% confidence in significant result

generate random data
calculate stats
pull p value
repeat 100x storing each in vector
calculate percent sig

change sample size up 1
repeat above 10-100 storing each in vector

graph sample size (x) versus likelihood of significant results (y)



ggsave(filename="Homework9RegressionPlot.pdf",
       plot=regPlot,
       device="pdf")